A Parallel General Purpose Multi-Objective Optimization Framework, with Application To Electron Beam Dynamics

### 2019-02-01 paper/submission/theor-experiment.tex
lot's of small edits, mostly commas

In this section a numerical results of the validation benchmark and
  optimization of a photoinjector operated in the space charge dominated regime is presented.

@@

To ensure that the optimizer works correctly the benchmark
  problem (\ref{eqn:bench}) was solved.
To that end, we use a metric for comparing the quality of a Pareto
  front.
Given a point in the Pareto set, we compute the $m$ dimensional volume (for
  $m$ objectives) of the dominated space, relative to a chosen origin.
This is visualized for $2$ objectives in Fig.~\ref{fig:hypervolume}.
For further information and details of the implementation see~\cite{whbb:12}.
Figure~\ref{fig:pisa_bench} and the corresponding hypervolume values in
  Table~\ref{tbl:bench_rms_error}, show expected convergence.
The reference Pareto front is clearly very well approximated.
It took a total of 1100 function evaluations to perform this computation.
The hypervolume of the reference solution ($0.6575$) for our benchmark was
  computed by sampling the solution provided in~\cite{hbwh:05}.

@@

\begin{table}%[h!]
\begin{center}
  \caption{Convergence of benchmark problem with errors relative to
    hypervolume of sampled reference solution.}
  \label{tbl:bench_rms_error}
  \begin{tabular}{c|c|c}
    \hline\noalign{\smallskip}
    tot.\ function  & hypervolume & relative error\\
    evaluations    & & \\
    \noalign{\smallskip}\hline\noalign{\smallskip}
    100  &  0.859753 & $3.076 \times 10^{-1}$ \\
    200  &  0.784943 & $1.938 \times 10^{-1}$ \\
    500  &  0.685183 & $4.210 \times 10^{-2}$ \\
    900  &  0.661898 & $6.689 \times 10^{-3}$ \\
    1100 &  0.657615 & $1.749 \times 10^{-4}$ \\
    \noalign{\smallskip}\hline
  \end{tabular}
\end{center}
\end{table}

@@

\subsubsection{Time Step Scan} \label{awa:subsection:test}
Before running a full scale optimization of the problem described in Subsection \ref{awaproblem}, 
a study on time step and number of particles in the simulation model 
was done to reduce the time of the simulation while 
maintaining the physics of interest. 
The grid size $16 \times 16 \times 32$ was chosen, 
and parallelized in the x and y directions.
After comparing several options (1,000, 10,000, 20,000, 50,000, 100,000) 
with a small time step, the number of particles was fixed at 10,000.
Next several time steps were explored, see Table~\ref{timestep}.
The largest steps were too big to resolve the beam parameters accurately.
See low fidelity plot in Fig.~\ref{tstep} for $dT=$\num{5E-11} results.

\begin{table}%[h!]
	\begin{center}
		\caption{Checkmarks (\cmark) indicate desired beam parameters are resolved at that time step. 
			An (\xmark) indicates the time step is too large, and results are nonphysical.}
		\label{timestep}
		\begin{tabular*}{0.48\textwidth}{c|c|c|c}
			\hline\noalign{\smallskip}
			Time Step, dT (s) \quad & Linac & Drift & Quadrupoles \\
			\noalign{\smallskip}\hline\noalign{\smallskip}
			$5 \times10^{-10}$  & \xmark & \xmark & \xmark \\
			$1 \times10^{-10}$  & \xmark & \xmark & \xmark \\
			$5 \times10^{-11}$  & \xmark & \xmark & \xmark \\
			$1 \times10^{-11}$  & \cmark & \cmark & \xmark \\
			$5 \times10^{-12}$  & \cmark & \cmark & \xmark \\
			$1 \times10^{-12}$  & \cmark & \cmark & \cmark \\
			\noalign{\smallskip}\hline
		\end{tabular*}
	\end{center}
\end{table}

@@

\begin{figure}%[h]
	\centering
	\includegraphics[width=\linewidth]{figures/timestep_comparison_paper}
	\caption{Comparison of different fidelity models (dT stands for time step).}
	\label{tstep}
\end{figure}

\subsubsection{Hyper parameter Scan}
While the optimization problem and goals were well defined (Subsection \ref{awaproblem}), 
it was not clear what the best hyper parameters for the genetic algorithm would be.
These parameters include gene mutation probability, mutation probability, 
recombination probability, number of individuals, 
and number of generations to complete. 
Given the beam line in Fig.~\ref{awa-linac},
four small optimization experiments were done with various hyper parameters. 
Similar to the time step scan, 
the goal of this exercise was to determine which set of optimization
parameters strongly influence the results, 
and whether there was a time to solution difference.
From here on, we will reference each experiment as ex-1, ex-2, ex-3, and ex-4
as shown in Table \ref{extable}. 
\begin{table}%[h!]
	\begin{center}
		\caption{Input Parameters for initial twenty four hour AWA optimization experiments. 
			The gene mutation probability was equal to the mutation probability (not shown) in all four experiments. 
			The max number of individuals per generation was~80.}
		\label{extable}
		\begin{tabular*}{0.48\textwidth}{l c|c|c}
			& Gene Mutation \qquad& Recombination \qquad& Number of completed\vspace{-0.75em}\\
			& Probability \qquad& Probability \qquad&  generations \\
			\midrule
			ex-1 &  0.1  & 0.9  &  96 \\
			ex-2 &  0.3  & 0.7  &  81 \\
			ex-3 &  0.8  & 0.2  &  53 \\
			ex-4 &  0.01 & 0.09 &  95 \\
		\end{tabular*}
	\end{center}
\end{table}

@@

\begin{figure}
	\includegraphics[width=0.65\textwidth]{figures/xy-max-min-sigma-paper}
	\caption{Optimized beam sizes along high charge beam line. The gun is located at $s=0$, 
	both x and y beam sizes are shown. The black line represents the relevant beam line aperture, while
	the green line indicates the location of the optimization.}
	\label{fig:stat}
\end{figure}
\begin{table}%[h!]
	\begin{center}
		\caption{Input Parameters for the optimized solution shown in Figures~\ref{fig:pareto1} and \ref{fig:stat}.}
		\label{tab:designopt}   
		\begin{tabular}{lcc}
			\toprule
			\textbf{Design Variable} \qquad \qquad & \textbf{Unit}	\quad&  \textbf{Value}  \\ 
			\midrule
			{Buck Focusing Solenoid} \qquad \qquad & amps	\quad & 478 \\
			Matching Solenoid &	amps	& 197	  \\
			Quadrupole 1& T-m		& -0.8	\\ 
			Quadrupole 2& T-m		& 0.9	\\
			Quadrupole 3 & T-m		& 0.8	\\
			Quadrupole 4 & T-m		& -1.0	\\ 
			Bunch Length & mm 		& 1.5	\\
			\bottomrule
		\end{tabular}
	\end{center}
\end{table}

>>>>

In this section numerical results of the validation benchmark and optimization 
of a photoinjector operated in the space charge dominated regime is presented.

@@

To ensure that the optimizer works correctly, the benchmark
  problem (\ref{eqn:bench}) was solved.
To that end, we use a metric for comparing the quality of a Pareto front.
Given a point in the Pareto set, we compute the $m$ dimensional volume (for
  $m$ objectives) of the dominated space, relative to a chosen origin.
This is visualized for $2$ objectives in Fig.~\ref{fig:hypervolume}.
For further information and details of the implementation see~\cite{whbb:12}.
Figure~\ref{fig:pisa_bench} and the corresponding hypervolume values in
  Table~\ref{tbl:bench_rms_error} show expected convergence.
The reference Pareto front is clearly very well approximated.
It took a total of 1100 function evaluations to perform this computation.
The hypervolume of the reference solution ($0.6575$) for our benchmark was
  computed by sampling the solution provided in~\cite{hbwh:05}.

@@

\begin{table}%[h!]
\begin{center}
  \caption{Convergence of benchmark problem with errors relative to
    hypervolume of sampled reference solution.}
  \label{tbl:bench_rms_error}
  \begin{tabular}{c|c|c}
	\cline{1-3}
    tot.\ function  & hypervolume & relative error\\
    evaluations    & & \\
	\cline{1-3}
    100  &  0.859753 & $3.076 \times 10^{-1}$ \\
    200  &  0.784943 & $1.938 \times 10^{-1}$ \\
    500  &  0.685183 & $4.210 \times 10^{-2}$ \\
    900  &  0.661898 & $6.689 \times 10^{-3}$ \\
    1100 &  0.657615 & $1.749 \times 10^{-4}$ \\
	\cline{1-3}
  \end{tabular}
\end{center}
\end{table}

@@

\subsubsection{Time Step Scan} \label{awa:subsection:test}
Before running a full scale optimization of the problem described in Subsection \ref{awaproblem}, 
a study on time step and number of particles in the simulation model 
was done to reduce the time of the simulation while 
maintaining the physics of interest. 
The grid size $16 \times 16 \times 32$ was chosen, 
and parallelized in the x and y directions.
After comparing several options (1,000, 10,000, 20,000, 50,000, 100,000) 
with a small time step, the number of particles was fixed at 10,000.
Next several time steps were explored, see Table~\ref{timestep}.
The largest steps were too big to resolve the beam parameters accurately.
See low fidelity plot in Fig.~\ref{tstep} for $dT=$\num{5E-11} results.  
\begin{table}%[h!]
	\begin{center}
		\caption{Checkmarks (\cmark) indicate desired beam parameters are resolved at that time step. 
			An (\xmark) indicates the time step is too large, and results are nonphysical.}
		\label{timestep}
		\begin{tabular*}{0.48\textwidth}{c|c|c|c}
	\cline{1-4}
			Time Step, dT (s) \quad & Linac & Drift & Quadrupoles \\
	\cline{1-4}
			$5 \times10^{-10}$  & \xmark & \xmark & \xmark \\
			$1 \times10^{-10}$  & \xmark & \xmark & \xmark \\
			$5 \times10^{-11}$  & \xmark & \xmark & \xmark \\
			$1 \times10^{-11}$  & \cmark & \cmark & \xmark \\
			$5 \times10^{-12}$  & \cmark & \cmark & \xmark \\
			$1 \times10^{-12}$  & \cmark & \cmark & \cmark \\
	\cline{1-4}
		\end{tabular*}
	\end{center}
\end{table}

@@

\begin{figure}%[h]
	\centering
	\includegraphics[width=0.8\linewidth]{figures/timestep_comparison_paper}
	\caption{Comparison of different fidelity models (dT stands for time step).}
	\label{tstep}
\end{figure}

\subsubsection{Hyper parameter Scan}
While the optimization problem and goals were well defined (Subsection \ref{awaproblem}), 
it was not clear what the best hyper parameters for the genetic algorithm would be.
These parameters include gene mutation probability, mutation probability, 
recombination probability, number of individuals, 
and number of generations to complete. 
Given the beam line in Fig.~\ref{awa-linac},
four small optimization experiments were done with various hyper parameters. 
Similar to the time step scan, 
the goal of this exercise was to determine which set of optimization
parameters strongly influence the results, 
and whether there was a time to solution difference.
From here on, we will reference each experiment as ex-1, ex-2, ex-3, and ex-4
as shown in Table \ref{extable}. 
\begin{table}%[h!]
\begin{center}
\caption{Input Parameters for initial twenty four hour AWA optimization experiments. 
The gene mutation probability was equal to the mutation probability (not shown) in all four experiments. 
The max number of individuals per generation was~80.}
\label{extable}
\begin{tabular}{c|c|c|c}
\cline{1-4}
& Gene Mutation \qquad& Recombination \qquad & Number of completed \vspace{-0.75em}\\
Experiment & Probability & Probability &  generations \\ 
\cline{1-4}
ex-1 &  0.1  & 0.9  &  96 \\
ex-2 &  0.3  & 0.7  &  81 \\
ex-3 &  0.8  & 0.2  &  53 \\
ex-4 &  0.01 & 0.09 &  95 \\ 
\cline{1-4}
\end{tabular}
\end{center}
\end{table}

@@

\begin{figure}
	\includegraphics[width=0.65\textwidth]{figures/xy-max-min-sigma-paper}
	\caption{Optimized beam sizes along high charge beam line. The gun is located at $s=0$, 
	both x and y beam sizes are shown. The black line represents the relevant beam line aperture, while
	the green line indicates the location of the optimization.}
	\label{fig:stat}
\end{figure}
\begin{table}%[h!]
	\begin{center}
		\caption{Input Parameters for the optimized solution shown in Figures~\ref{fig:pareto1} and \ref{fig:stat}.}
		\label{tab:designopt}   
		\begin{tabular}{l|c|c}
			\cline{1-3}
			\textbf{Design Variable} \qquad \qquad & \textbf{Unit}	\quad&  \textbf{Value}  \\ 
			\cline{1-3}
			{Buck Focusing Solenoid} \qquad \qquad & amps	\quad & 478 \\
			Matching Solenoid &	amps	& 197	  \\
			Quadrupole 1& T-m		& -0.8	\\ 
			Quadrupole 2& T-m		& 0.9	\\
			Quadrupole 3 & T-m		& 0.8	\\
			Quadrupole 4 & T-m		& -1.0	\\ 
			Bunch Length & mm 		& 1.5	\\
			\cline{1-3}
		\end{tabular}
	\end{center}
\end{table}
